---
title: "Exploratory_analysis"
author: "Clara Mugnai"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
library(dplyr)
library(lme4)
turnout_data <- read_excel("SYE data sheet.xlsx",
 col_types = c(
 ))
```
Data summary:
  "Horse" - Horse name
  "Time_observed(minutes)" - Time horse was observed
  "Time_of_day" - Window of time horse was observed
  "Age" - Age of horse
  "Personality" - Personality type of specific horse
  "Date" - Date horse was observed
  "Flymask" - Whether or not horse was wearing a flymask
  "Flysheet" - Whether or not horse was wearing a flysheet
  "Group" - Whether or not horse was in a group
  "Type_of_turnout" - Which turnout field the horse was in for the time observed
  "Frequency_of_tail_swishes" - how often the horse swished its tail during turnout
  "Minutes_grazing" - How many minutes of turnout the horse spent grazing
  "Dialated_nostrils" - How many times the horse dialated its nostrils during turnout
  "Rigid_body_posture" - How many times the horse made its body rigid during turnout
  "minutes_spent_pacing" - Haw many minutes of turnout the horse spent pacing or stressed 
  "number_whinnies" - How many times the horse whinnied in turnout
  "number_positive_interactions_w/friends" - How many positive interactions the horse experienced with other horses
  "number_negative_interactions_w/friends" - How many negative interactions the horse experienced with other horses
  "time_layingdown" - Minutes the horse spent laying down

```{r}
#Using total negative signs instead of Overall because that is not as biased towards horses with turnout where there are grass.
turnout_data %>% group_by(Horse) %>% summarise(mean(Totalnegativesigns))
turnout_data %>% group_by(Personality) %>% summarise(mean(Totalnegativesigns))
ggplot(data = turnout_data, mapping = aes(x = Personality, y = Totalnegativesigns)) +
  geom_boxplot()
#interesting but have to watch how many in each group there are
ggplot(data = turnout_data, mapping = aes(x = Time_of_day, y = Totalnegativesigns)) + geom_boxplot()
# chaotic
ggplot(data = turnout_data, mapping = aes(x = Age, y = Totalnegativesigns)) + geom_point() + geom_smooth(se = FALSE)
#possibly significant
ggplot(data = turnout_data, mapping = aes(x = Time_of_day, y = Totalnegativesigns)) + geom_point() + geom_smooth(se = FALSE)
# Why am I not getting a smoother for this plot?

ggplot(data = turnout_data, mapping = aes(x = Flymask, y = Totalnegativesigns)) + geom_boxplot()
#possibly significant
ggplot(data = turnout_data, mapping = aes(x = Flysheet, y = Totalnegativesigns)) + geom_boxplot()
# not important

ggplot(data = turnout_data, mapping = aes(x = Type_of_turnout, y = Totalnegativesigns)) + geom_boxplot()
# interesting but need to consider other causes for the trends.

ggplot(turnout_data, aes(x=minutes_spent_pacing,y=Totalnegativesigns)) + 
  geom_point() + 
  geom_smooth(method="lm") +
  facet_wrap("Horse")

ggplot(data=turnout_data,aes(x=Age,y=Totalnegativesigns,color=Group)) + 
  geom_point() + 
  geom_smooth(method="lm")

ggplot(data=turnout_data,aes(x=Age,y=Totalnegativesigns,color=Flymask)) + 
  geom_point() + 
  geom_smooth(method="lm")

ggplot(turnout_data, aes(x=Personality,y=Totalnegativesigns)) + 
  geom_boxplot(fill="steelblue")

ggplot(data=turnout_data,aes(x=Age,y=Totalnegativesigns,color=Personality)) + 
  geom_point() + 
  geom_smooth(method="lm") 
```

- Exploring multilevel scenarios, it looks like Age and Group may be related, with different slopes for group and non-group, so as age increases in a group the group does better but as age increases in individual horses they do worse. 

- Flymask also seems to have significant differences in slopes and intercepts. 


- The different personalities definitely are different, and we have seen that each horse is different so we do need to account for the grouped horses. 

- the personalities seem to make a difference in the slopes here but we have to be mindful of the fact that there are very few of each personality. 


Unconditional means model:
```{r}
unmeanmod = lmer(Totalnegativesigns ~ 1 + (1|Horse), data = turnout_data, REML = T)
summary(unmeanmod)
```
- looking at the unconditional means model, the ICC value is (35.79/(16.58+35.79) = 0.68, so this is a multilevel data set and needs to be allowed to have different intercepts and "slopes" for different horses. 
